---
title: "megadepth vs pyBigWig vs rtracklayer"
author: "Leo Collado-Torres, David Zhang, Chris Wilks"
output: 
  github_document
---

```{r setup, include = FALSE}

library(magrittr)

knitr::opts_chunk$set(echo = T, message = F, warning = T)

```

> Aim: Check the speed of different tools for loading coverage.

# Background

There are multiple tools that are able to obtain coverage from BigWig/BAM files. Specifically, here we test the tools [megadepth](https://github.com/ChristopherWilks/megadepth), [rtracklayer](https://bioconductor.org/packages/release/bioc/html/rtracklayer.html) and [pyBigWig](https://github.com/deeptools/pyBigWig). Some operations requiring coverage can be very computationally intensive - for example, the [dasper](https://github.com/dzhang32/dasper) R package requires loading coverage across 3 regions per junction per sample. Here, we aim to find the fastest tool for loading coverage from BigWig files by comparing the speed of the `megadepth` vs `rtracklayer` vs `pyBigwig`.

# Input data

First off, we obtain paths to the annotation ranges and BigWig files to load coverage from. For these, we have used the examples hosted on https://github.com/ChristopherWilks/megadepth. 

## Annotation co-ordinates

- We'll work with the example set of ranges found [here](https://github.com/ChristopherWilks/megadepth/blob/master/tests/testbw2.bed). 

```{r}

annotation_file <- system.file("tests", "testbw2.bed", package = "megadepth", mustWork = TRUE)

# convert to a GRanges for rtracklayer
annotation <- annotation_file %>% 
  rtracklayer::import(format = "BED")

annotation

# check ranges are identical
annotation_check <- annotation_file %>% 
  readr::read_delim("\t", col_names = c("chr", "start", "end")) %>% 
  GenomicRanges::GRanges()

# change BED (0-based coords) to match GRanges
GenomicRanges::start(annotation_check) <- GenomicRanges::start(annotation_check) + 1 

testthat::expect_identical(annotation_check, annotation)

```

## BigWig files

- We will obtain the paths to the small example BigWig as part of the `megadepth` R package. 

```{r}

bw_local <- system.file("tests", "test.bam.all.bw", package = "megadepth", mustWork = TRUE)
bw_remote <- "https://github.com/LieberInstitute/megadepth/raw/master/inst/tests/test.bam.all.bw"

```

# Are coverage values obtained through the different methods consistent?

- First, I will check whether the coverage values obtained through `rtracklayer` match those obtained through `megadepth`.  

```{r}

# get base-wise coverage as numeric list for summarisation after
rt_cov <- rtracklayer::import.bw(con = bw_local, 
                                 which = annotation, 
                                 as = "NumericList")

# rtracklayer::import.bw seems to muddle the output order
# reorder - CHECK WHY 
rt_cov <- rt_cov[c(2, 3, 4, 1)] %>% BiocGenerics::mean()

md_cov <- megadepth::get_coverage(bigwig_file = bw_local, 
                                  annotation = annotation_file, 
                                  op = "mean")

try(testthat::expect_equivalent(rt_cov %>% round(3), 
                                md_cov$cov))

```

# Test speed 

- Here, we will test the speed of `rtracklayer`, `megadepth` and `pyBigwig` when obtaining coverage from a BigWig file. To make sure the results are comparable, we will use the identical set of ranges and BigWigs initialised above. 

## rtracklayer & megadepth

- `rtracklayer` and `megadepth` can be easily called in R therefore for testing the runtime of these tools I will use the R package [microbenchmark](https://cran.r-project.org/web/packages/microbenchmark/index.html).

```{r, message = FALSE}

n_iters <- 5

benchmark_rt_md <- function(bw, times = 10){
  
  runtime <- microbenchmark::microbenchmark(
    "rt"= { 
      rtracklayer::import.bw(con = bw, 
                             which = annotation, 
                             as = "NumericList")
      }, 
    "md" = {
      megadepth::get_coverage(bigwig_file = bw, 
                              annotation = annotation_file, 
                              op = "mean")
    }, 
    times = times)
  
  runtime <- runtime %>% 
    dplyr::mutate(time = time / 1e9) %>% # convert time to secs
    dplyr::rename(tool = expr, run_time = time)
  
  return(runtime)
  
}

runtime_rt_md <- rbind(benchmark_rt_md(bw = bw_local, times = n_iters) %>% 
                         dplyr::mutate(local_remote = "local"), 
                       benchmark_rt_md(bw = bw_remote, times = n_iters) %>% 
                         dplyr::mutate(local_remote = "remote"))

```

## pyBigWig {.tabset}

- As `pyBigWig` is a python package, to ensure the paths to the BigWig and annotation ranges remain the same, I will call the script found `r here::here("analysis", "test_pyBigWig.py")`. This creates a function that calls `pyBigWig` for a set number of iterations and saves the runtime for each.
- To run this, you must have python3.X available on your OS with the modules `pyBigWig`, `pandas` and `numpy` installed. 

```{r}

benchmark_pybw <- function(bw, times = 10){
  
  out_path <- paste0(tempdir(), "/pyBigWig_run_time.txt")
  
  system2("python3", 
          args = c(here::here("analysis", "test_pyBigWig.py"),
                   bw, 
                   annotation_file, 
                   out_path, 
                   times)
  )
  
  runtime <- readr::read_delim(out_path, delim = "\t")
  
  return(runtime)
  
}

runtime_pybw <- rbind(benchmark_pybw(bw = bw_local, times = n_iters) %>% 
                         dplyr::mutate(local_remote = "local"), 
                      benchmark_pybw(bw = bw_remote, times = n_iters) %>% 
                         dplyr::mutate(local_remote = "remote"))

```

# Compare speeds 

```{r, eval = FALSE}

comp_time <- runtime_rt_md %>% 
  as.data.frame() %>% 
  rbind(runtime_pybw) %>% 
  dplyr::mutate(tool = tool %>% factor(c("md", "rt", "pyBigwig")))

comp_time_plot <- 
  ggplot2::ggplot(comp_time, ggplot2::aes(x = tool, 
                                          y = run_time, 
                                          colour = local_remote)) + 
  ggplot2::geom_boxplot() +  
  ggplot2::geom_jitter() + 
  ggplot2::scale_x_discrete(name = "Tool", 
                            labels = c("megadepth", 
                                       "rtracklayer", 
                                       "pyBigWig")) + 
  ggplot2::scale_y_continuous(name = paste0("Run time for ", length(annotation), " regions (seconds)")) + 
  ggplot2::scale_colour_manual(values = ggpubr::get_palette("npg", 2), 
                               guide = "none") + 
  ggplot2::facet_wrap(~ local_remote, 
                      scales = "free") + 
  ggplot2::theme_bw(base_size = 20) 

comp_time_plot %>% 
  ggplot2::ggsave(filename = "md_rt_pybw_runtime.png", 
                  plot = ., 
                  path = here::here("analysis"), 
                  dpi = 300, 
                  width = 10, height = 5)
  
```

# Reproducibility

```{r echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```


